\LARGE{\bf{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}}\footnotemark

### 2014-09-04 ResponseToReview.tex

\res{We did not measure ear architecture or internode length. 
While \emph{tb1} does impact these traits, our primary interest was in the effect of the \emph{Hopscotch} insertion at \emph{tb1} on tillering given that this is an ecologically relevant phenotype that is known to be substantially affected by \emph{Hopscotch} in a maize background.
Both additional greenhouse and expression analyses would indeed be useful avenues of further investigation.
Nonetheless, these are extensive additional experiments which we are unable to include at this time. 
We have, however, added text to the conclusion (Page 15, lines 372-377) of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte:
``Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g. \emph{gt1}, \emph{tru1}, and \emph{te1}).  These data, in conjunction with more exhaustive phenotyping, should help to further clarify the ecological significance of the domesticated \emph{tb1} allele in natural populations of teosinte."}

One minor issue: in Fig. 1, it would be nice to see where the Balsas River Basin is in relation to the sampling shown.\\

\res{We have added the Balsas River in a revised version of Figure 1.}

\section*{Response to Reviewer 2} 
5) Although this paper brings up an interesting question regarding the role of tb1 standing variation in teosinte, it is not able to answer it completely.  It presents results that have been published previously about the presence of hopscotch in teosinte, maize, and mexicana.\\

@@

6) This paper is fairly maize-centric. A more general message is included in the discussion, not as much in the introduction, and not at all in the abstract. This could be easily remedied.\\

\lev{need to add something here. I had added something I thought in the text...}

Methods and statistics were appropriately used as far as my expertise allows me to judge.
Specific suggestions follow:

7) Introduction:
Page 4, Lines 7 - 9: You say ``Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovachMcCouch2008}.".  Can you tell us why an understanding is useful? Can you give some examples?\\

\res{We have expanded on this concept in the introduction on page \lev{x} lines \lev{x-x} ``Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \citep{Ross-Ibarra2007}. 
Plant domestication usually involves a suite of phenotypic changes such as loss of seed shattering and increased fruit or grain size, which are commonly referred to as `domestication syndrome' \citep{OlsenWendel2013}, and much of the study of domestication has focused on understanding the genetic variation underlying these traits \citep{OlsenGross2010}. 
Because most domesticates show reduced genetic diversity relative to their wild counterparts, effort has been made to identify agronomically useful variation in crop wild relatives \citep{flint2009wide}. Often, after identification, the alleles conferring these beneficial traits are bred into domesticates for crop improvement. For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovackMcCouch2008}. In addition to researching the role of wild alleles in domesticates, researchers have also investigated the role of variation in domesticated taxa in the evolution of feral and weedy populations \citep{ellstrand2010crops}. A recent paper examining the genomic changes associated with rabbit domestication found that many domestication phenotypes are due to mutations in many small effect loci, and that many of these mutations are still found segregating in natural populations of wild rabbit, suggesting that selection acting on standing variation \citep{Rabbit}. There are numerous examples of weedy plants evolving to mimic their domestic counterpart, and even some instances where weedy plants have evolved domesticates (reviewed in \citealt{ellstrand2010crops}). But even though domesticated alleles are often found segregating in wild relatives \citep{Gallavotti2004, SigmonVollbrecht2010}, we know almost nothing about the ecological role of this variation in natural populations. In this paper we present an ecological genetic analysis of the domestication locus \emph{tb1} in natural populations of the wild ancestor of domesticated maize."}

@@

17) Page 9 lines 2 - 3: ``Plants were watered three times a day by hand and with an automatic drip containing 10-20-10 fertilizer." This sounds odd; by hand or automatically? Or you turned on the drip by hand? In which case, it was not automatic\\

\res{We have edited the text to clarify, page 9, lines 199-200 : ``Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer, which was supplemented with hand watering on extremely hot and dry days."}

18) Page 9 lines 7 - 8: ``Culm diameter is not believed to be correlated with tillering index or variation at tb1." So, why did you do it? Because of the last sentence in the materials and methods? Perhaps do not mention here that is not correlated; wait until the end of the materials and methods where you seem to explain it.\\

\res{Thank you for the suggestion, we have edited the text accordingly and have added our explanation of the inclusion of culm diameter at the end of the Materials and Methods section, \lev{page x, linesx} : ``Additionally, in order to compare any association between \emph{Hopscotch} genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS. Culm diameter is not believed to be correlated with tillering index or variation at \emph{tb1} and is used as our independent trait for phenotyping analyses."}

### 2014-09-04 manuscript.tex
edits
Edits to response to review, references, introduction rewrite

Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \citep{Ross-Ibarra2007}. 
Plant domestication usually involves a suite of phenotypic changes such as loss of seed shattering and increased fruit or grain size, which are commonly referred to as `domestication syndrome' \citep{OlsenWendel2013}, and much of the study of domestication has focused on understanding the genetic variation underlying these traits \citep{OlsenGross2010}. 
Because most domesticates show reduced genetic diversity relative to their wild counterparts, effort has been made to identify agronomically useful variation in crop wild relatives \citep{flint2009wide} \jri{add some more}. \mbh{this reference should be a gold mine: \url{http://bioscience.oxfordjournals.org/content/61/7/559.full}}
Researchers have also investigated the role of variation in domesticated taxa in the evolution of feral and weedy populations \citep{ellstrand2010crops} \jri{add more}. \mbh{It looks like the Ellstrand paper is a review; if so and there are a number of good examples of this point, I would simply say ``(reviewed in \citealt{ellstrand2010crops})"}But even though domesticated alleles are often found segregating in wild relatives \citep{Gallavotti2004, SigmonVollbrecht2010}, we know almost nothing about the ecological role of this variation in natural populations. \jri{add some examples, can include teosinte ones mentioned below but also a few other crops} \mbh{I added some maize/teosinte examples...agree with Jeff that examples from other species would be good...here's a very recent example from a rabbit domestication paper: \url{http://www.sciencemag.org/content/345/6200/1074.full}}
In this paper we present an ecological genetic analysis of the domestication locus \emph{tb1} in natural populations of the wild ancestor of domesticated maize. 
%some effort has been invested into identifying agronomically useful loci in wild relatives \citep{}understanding how alleles from wild relatives behave in domesticated crops \citep[e.g.][]{Bai2007}, or identifying agronomically usefulvery little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.
%In addition to the loss of diversity due to selection on loci important for domestication, demography and the effects of linked selection have acted to decrease diversity genome-wide, leading to lower genetic diversity in many domesticates relative to their wild progenitors \citep{KovachMcCouch2008}.

@@

In conclusion, our findings demonstrate that the \emph{Hopscotch} allele is widespread in populations of \emph{parviglumis} and \emph{mexicana} and occasionally at high allele frequencies. 
Analysis of linkage using SNPs from across chromosome 1 does not suggest that the \emph{Hopscotch} allele is present in these populations due to recent introgression, and it seems unlikely that the insertion would have drifted to high frequency in multiple populations. 
We do, however, find preliminary evidence of selection on the \emph{tb1} locus in \emph{parviglumis}.  Coupled with our observation of high frequency of the \emph{Hopscotch} insertion in a number of populations, this suggests that the locus may play an ecological role in teosinte. 
In contrast to domesticated maize, the \emph{Hopscotch} insertion does not appear to have a large effect on tillering in a diverse sample of \emph{parviglumis} from a natural population and the phenotypic consequences of variation at \emph{tb1} thus remain unclear. Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g. \emph{gt1}, \emph{tru1}, and \emph{te1}) as well as include a more exhaustive phenotyping including all traits.  nce of the domesticated \emph{tb1} allele in natural populations of teosinte.

>>>>

Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \citep{Ross-Ibarra2007}. 
Plant domestication usually involves a suite of phenotypic changes such as loss of seed shattering and increased fruit or grain size, which are commonly referred to as `domestication syndrome' \citep{OlsenWendel2013}, and much of the study of domestication has focused on understanding the genetic variation underlying these traits \citep{OlsenGross2010}. 
Because most domesticates show reduced genetic diversity relative to their wild counterparts, effort has been made to identify agronomically useful variation in crop wild relatives \citep{flint2009wide}. Often, after identification, the alleles conferring these beneficial traits are bred into domesticates for crop improvement. For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovackMcCouch2008}. In addition to researching the role of wild alleles in domesticates, researchers have also investigated the role of variation in domesticated taxa in the evolution of feral and weedy populations \citep{ellstrand2010crops}. A recent paper examining the genomic changes associated with rabbit domestication found that many domestication phenotypes are due to mutations in many small effect loci, and that many of these mutations are still found segregating in natural populations of wild rabbit, suggesting that selection acting on standing variation \citep{Rabbit}. There are numerous examples of weedy plants evolving to mimic their domestic counterpart, and even some instances where weedy plants have evolved domesticates (reviewed in \citealt{ellstrand2010crops}). But even though domesticated alleles are often found segregating in wild relatives \citep{Gallavotti2004, SigmonVollbrecht2010}, we know almost nothing about the ecological role of this variation in natural populations. In this paper we present an ecological genetic analysis of the domestication locus \emph{tb1} in natural populations of the wild ancestor of domesticated maize. 
%some effort has been invested into identifying agronomically useful loci in wild relatives \citep{}understanding how alleles from wild relatives behave in domesticated crops \citep[e.g.][]{Bai2007}, or identifying agronomically usefulvery little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.
%In addition to the loss of diversity due to selection on loci important for domestication, demography and the effects of linked selection have acted to decrease diversity genome-wide, leading to lower genetic diversity in many domesticates relative to their wild progenitors \citep{KovachMcCouch2008}.

@@

In conclusion, our findings demonstrate that the \emph{Hopscotch} allele is widespread in populations of \emph{parviglumis} and \emph{mexicana} and occasionally at high allele frequencies. 
Analysis of linkage using SNPs from across chromosome 1 does not suggest that the \emph{Hopscotch} allele is present in these populations due to recent introgression, and it seems unlikely that the insertion would have drifted to high frequency in multiple populations. 
We do, however, find preliminary evidence of selection on the \emph{tb1} locus in \emph{parviglumis}.  Coupled with our observation of high frequency of the \emph{Hopscotch} insertion in a number of populations, this suggests that the locus may play an ecological role in teosinte. 
In contrast to domesticated maize, the \emph{Hopscotch} insertion does not appear to have a large effect on tillering in a diverse sample of \emph{parviglumis} from a natural population and the phenotypic consequences of variation at \emph{tb1} thus remain unclear. Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g. \emph{gt1}, \emph{tru1}, and \emph{te1}). These data, in conjunction with more exhaustive phenotyping, should help to further clarify the ecological significance of the domesticated \emph{tb1} allele in natural populations of teosinte.
